Loss recovery and tail loss performance in Multipath TCP

### 2016-11-21 main.tex
update

Through this set of preliminary experiments, we would like to understand the behavior of MPTCP with and without TLP and ER by monitoring burst completion time, congestion window evolution and timeout 
behavior.
\section{Related Work}\label{relwork}

\section{Scope}\label{scope}

We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation. As per the specification of Multipath TCP introduced in 
MPTCP features acknowledgements at connection-level as well as
   subflow-level acknowledgements, in order to provide a robust service
   to the application.

@@

Planned experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
Client is connected to two wireless interfaces 3G and WLAN. Server is connected to a wired router.
Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
Traffic generation is planned with iperf, netperf and d-itg as different runs.

@@

      \textbf{Run Length} & \multicolumn{10}{c|}{10min}\\
      \hline 	
      \textbf{RTT} & \multicolumn{5}{c|}{200ms(3G)} & \multicolumn{5}{c|}{20ms(WLAN)} \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{5}{c|}{7Mbps(3G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
      \hline
      \textbf{Loss} & \multicolumn{5}{c|}{0(3G)} & \multicolumn{5}{c|}{2\%(WLAN)} \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Random}\\
      \hline
\end{tabular}
\caption{Connection parameters}\label{tab1}
\end{center}
\end{table}
\end{center}

\subsection{Testing with Deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses generated by using associating netem with corresponding interfaces and dropping the packets(Cite KAU paper).\ref{timing}\ref{1p}\ref{1pp}\ref{2p}

\section{Observations and Discussion}\label{disc}
Preliminary experiments on CORE emulator with the setup mentioned in section\ref{exsetup} showed several cases where the behavior on losses differ.

@@

@@

@@

\section{Tail loss scenarios}
We tried to understand the behavior of MPTCP retransmissions at different flow levels in the following tail loss scenarios~\ref{ret3}.
MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
is required to understand the losses at the subflow level.

@@

The linux sysctl setting tcp\_early\_retrans is 3 by default. This option enables both Early retransmit and TLP. 
Further testcases should be considered with options 0 and 4.

\section{Improvements to TLP}
The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occuring due to tail losses (losses at end of transactions) into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout.

>>>>

Through this set of preliminary experiments, we would like to understand the behavior of MPTCP with and without TLP and ER by monitoring burst completion time, congestion window evolution and timeout 
behavior.

\section{Related Work}\label{relwork}
TBD.
\section{Scope}\label{scope}

We try to understand the loss recovery and retransmission policies in state of the art MPTCP linux implementation version 0.91. As per the specification of Multipath TCP introduced in 
MPTCP features acknowledgements at connection-level as well as subflow-level, in order to provide a robust transport service to the application.

@@

Experiments use CORE emulation platform with a simple topology as depicted in Figure~\ref{fig1}.
Client is connected to two wireless interfaces 3G/4G and WLAN. Server is connected to a wired router.
Characteristics of the connection setup and assumptions about the parameters are provided in table~\ref{tab1}.
There is no attempt in this study to focus on the effect of link characteristics in retransmission performance.

@@

      \textbf{RTT} & \multicolumn{5}{c|}{20ms-120ms(3G/4G)} & \multicolumn{5}{c|}{20ms-120ms(WLAN)} \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{5}{c|}{54Mbps(3G/4G)} & \multicolumn{5}{c|}{54Mbps(WLAN)} \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Deterministic}\\
      \hline
\end{tabular}
\caption{Connection parameters}\label{tab1}
\end{center}
\end{table}
\end{center}

\subsection{Testing with Deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses generated by using associating netem with corresponding interfaces and dropping the packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions. We send two bursts of 80 
packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss. The expected retransmission behavior is shown in~\ref{timing}.

In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
ER and TLP. For this analysis, we used 0,3 and 4. In each setting, we calculate burst completion time for two burst of 80 packets each. The setup has server and client with MPTCP
enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.

@@

\section{Observations and Discussion}\label{disc}

Experiments performed on CORE emulator with the setup mentioned in section\ref{exsetup}.
MPTCP is sensitive to path asymmetry in general due to the default scheduler being shortest RTT based scheduler. Wireshark and tcptrace
analysis provides information on the TCP retransmissions at the meta level. mptcptrace provides information on flow sequence split at 
subflow level. Subflow level retransmissions are difficult to test from the output of mptcptrace analysis. Further deep packet inspection
required to understand the losses at the subflow level.

Figure~\ref{1p} provides burst completion time comparison of different ER sysctl settings considered with one packet loss. The results show the improvements
in burst completion time with ER and TLP enabled compared to both disabled only when there is delay difference. The retransmission of lost packet follows the behavior 
depicted in Figure~\ref{timing}.

@@

In packet trace inspection, it is easy to check the use of TLP with two packet tail loss. The last packet will be retransmitted first in this case.
With one packet tail loss one has to compare the RTT and last packet retransmission time to check the use of TLP.
To further validate the results with one packet tail loss, we consider two packet tail loss and the results are shown in~\ref{2p}.
In this case, ER=3 setting performs better than the other two settings in all scenarios except when the packets lost on high delay interface.

To further study the performance of TLP in improving latency, we tried to drop the probe packet along with the last packet. In this case ER=0
setting results much lower burst completion times.

@@

@@

\section{Improvements to TLP}
The goal of Tail Loss Probe(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occuring due to tail losses (losses at end of transactions) into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that current TLP implementation does not improve the performance in MPTCP and incur more latency.
